---
title: "US Murders Report"
author: "Leroy Buliro"
date: "2019 M10 15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a report on 2010 gun murder rates obtained from FBI reports. The original data was obtained from [this Wikipedia page](https://en.wikipedia.org/wiki/Murder_in_the_United_States_by_state).

We are going to use the following library:

```{r loading-libs, message=FALSE}
library(tidyverse)
```
  
Next we shall download the data and save it.

```{r}
url <- "https://raw.githubusercontent.com/rafalab/dslabs/master/inst/extdata/murders.csv"
dest_file <- "data/murders.csv"
download.file(url, destfile = dest_file)
```


then load and wrangle the data.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
murders <- read_csv("data/murders.csv")
murders <- murders %>% mutate(region=factor(region), rate=total/population * 10^5)
```

## Murder rate by state 

We note the large state to state variability by generating a barplot showing the murder rate by state:

```{r murder-rate-by-state, echo=FALSE}
murders %>% mutate(abb = reorder(abb, rate)) %>%
  ggplot(aes(abb, rate)) +
  geom_bar(width = 0.5, stat = "identity", color = "black") +
  coord_flip()
```